---
title: "Predicting EF from Cardiac MRIs: Descriptive Analysis and Simple Linear Regression"
author: "Shashin Chokshi and Jacob Albers"
date: "August 20, 2018"
output: rmarkdown::github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  fig.path = "README_figs/README-"
)
```

## Step 1: Ingest Data Processed by Python 

```{r, echo=TRUE, message=FALSE, warning=FALSE}

rm(list=ls())
input_data = read.csv("~/Code/uchicago_bmi/clinical_informatics/ef_prediction/kaggle_ndsb2/train_validate_preprocessing/train_enriched.csv")
test_data = read.csv("~/Code/uchicago_bmi/clinical_informatics/ef_prediction/kaggle_ndsb2/test_enriched.csv")

```

## Step 2: Load Key Libraries for Data Processing and Analysis

```{r, echo=TRUE, message=FALSE, warning=FALSE}

library(data.table)
library(Rmisc)
library(dplyr)
library(sqldf)
library(lubridate)
library(ggthemes)
library(ggplot2)
```

## Step 3: Process Data and Create Table One Summary 

TODO: Need to take out age

```{r, echo=TRUE, fig.height = 8, fig.width = 8}


train_data <- input_data %>%
              filter(Id <= 500) %>%
              select('patient_id', 'rows', 'columns', 'spacing', 'slice_thickness', 'plane', 'slice_count', 'up_down_agg', 'age_years', 'sex',
                       'small_slice_count', 'normal_slice_count', 'angle', 'Systole', 'Diastole') %>%
              mutate(ejection_fraction = round((Diastole-Systole)/Diastole * 100,1)) 

validate_data <- input_data %>%
                 filter(Id > 500) %>%
                select('patient_id', 'rows', 'columns', 'spacing', 'slice_thickness', 'plane', 'slice_count', 'up_down_agg', 'age_years', 'sex', 
                       'small_slice_count', 'normal_slice_count', 'angle', 'Systole', 'Diastole') %>%
                mutate(ejection_fraction = round((Diastole-Systole)/Diastole * 100,1)) 

test_data <- test_data %>%
                select('patient_id', 'rows', 'columns', 'spacing', 'slice_thickness', 'plane', 'slice_count', 'up_down_agg', 'age_years', 'sex', 
                       'small_slice_count', 'normal_slice_count', 'angle', 'Systole', 'Diastole') %>%
                mutate(ejection_fraction = round((Diastole-Systole)/Diastole * 100,1)) 


library(tableone)

dput(names(train_data))

myVars = c("patient_id", "rows", "columns", "spacing", "slice_thickness", 
"plane", "slice_count", "up_down_agg", "age_years", "sex", "small_slice_count", 
"normal_slice_count", "angle", "Systole", "Diastole", "ejection_fraction")

catVars = c("sex", "plane")

tab1 <- CreateTableOne(vars = myVars, data = train_data, factorVars = catVars)
print(tab1, showAllLevels = TRUE)

```

## Step 4: Visualize Jobs by Key Dimensions

```{r, echo=TRUE, message=FALSE, warning=FALSE, fig.height = 8, fig.width = 8}


# Systole Histogram
ggplot(train_data, aes(x=Systole)) + 
  geom_histogram(binwidth=5) +     
  labs(title="Systolic Volume", x = "End Systolic Volume (mLs)", y = "Number of Patients") + 
  theme_economist()

# Diastole Histogram
ggplot(train_data, aes(x=Diastole)) + 
  geom_histogram(binwidth=5) +     
  labs(title="Diastolic Volume", x = "End Diastolic Volume (mLs)", y = "Number of Patients") + 
  theme_economist()

# Slice Count Histogram
ggplot(train_data, aes(x=slice_count)) + 
  stat_count(binwidth=1) +     
  labs(title="Slice Count by Number of Patients", x = "Slice Count", y = "Number of Patients") + 
  theme_economist()

# Age vs. Sex
ggplot(train_data, aes(x=sex, y=age_years)) + 
  geom_boxplot() +     
  labs(title="Age vs Sex", x = "Sex", y = "Age (yrs)") + 
  theme_economist()


## Additional Graphs
ggplot(train_data, aes(x=age_years)) + 
  geom_histogram(binwidth=5) +     
  labs(title="Bandwidth by Count of Jobs", y = "Count of Jobs") + 
  theme_economist()


ggplot(train_data, aes(x=slice_thickness)) + 
  geom_histogram(binwidth=1) +     
  labs(title="Bandwidth by Count of Jobs", y = "Count of Jobs") + 
  theme_economist()

ggplot(train_data, aes(x=angle)) + 
  geom_histogram(binwidth=1) +     
  labs(title="Angle by Number of Patients", y = "Count of Jobs") + 
  theme_economist()


ggplot(train_data, aes(x=sex, y=age_years)) + 
  geom_boxplot() +     
  labs(title="Age for different Genders", x = "Age") + 
  theme_economist()





```


## Step 5: ## Step 5b: Run basic Linear Regression using DICOM metadata to Predict Values from Validate Dataset

```{r, echo=TRUE, message=FALSE, warning=FALSE,  fig.height = 8, fig.width = 8}

### Systolic Regression
sys_reg <- lm(Systole ~ patient_id + rows + columns + spacing + slice_thickness + plane + slice_count + up_down_agg + age_years + sex + small_slice_count + normal_slice_count + angle, data=train_data)
summary(sys_reg)

# Create Predicted Value
validate_data$PredictedSystole <-predict(sys_reg, newdata = validate_data)


# Plot Predicted vs Actual Systoliv Volume
ggplot(validate_data, aes(x=validate_data$Systole, y=validate_data$PredictedSystole)) + 
  geom_point()+
  labs(title="Predicted vs Actual Systolic Volume", x = "Predicted Volume (mL)", y = "Actual Volume (mL)") + 
  geom_smooth(method=lm)

# Calculate Key Metrics
cor(validate_data$Systole, validate_data$PredictedSystole)

# RMSE 
sqrt(sum((validate_data$Systole-validate_data$PredictedSystole)^2))
sqrt(sum((sys_reg$fitted.values-train_data$Systole)^2))

### Diastolic Regression

dias_reg <- glm(Diastole ~ patient_id + rows + columns + spacing + slice_thickness + plane + slice_count + up_down_agg + age_years + sex + small_slice_count + normal_slice_count + angle, data=train_data)
summary(dias_reg)

validate_data$PredictedDiastole <-predict(dias_reg, newdata = validate_data)

ggplot(validate_data, aes(x=validate_data$Diastole, y=validate_data$PredictedDiastole)) + 
  geom_point()+
  labs(title="Predicted vs Actual Diastolic Volume", x = "Predicted Volume (mL)", y = "Actual Volume (mL)") + 
  theme(plot.title = element_text(size=30)) + 
  geom_smooth(method=lm)

cor(validate_data$Diastole, validate_data$PredictedDiastole)
sqrt(sum((validate_data$Diastole-validate_data$PredictedDiastole)^2))


### Ejection Fraction
ef_reg <- glm(ejection_fraction ~ patient_id + rows + columns + spacing + slice_thickness + plane + slice_count + up_down_agg + age_years + sex + small_slice_count + normal_slice_count + angle, data=train_data)
summary(dias_reg)

validate_data$pred_ejection_fraction <-predict(ef_reg, newdata = validate_data)

ggplot(validate_data, aes(x=validate_data$ejection_fraction, y=validate_data$pred_ejection_fraction)) + 
  geom_point()+
  labs(title="Predicted vs Actual Ejection Fraction", x = "Predicted Ejection Fraction (%)", y = "Actual Ejection Fraction (%)") + 
  theme(plot.title = element_text(size=30)) + 
  geom_smooth(method=lm)

cor(validate_data$ejection_fraction, validate_data$pred_ejection_fraction)
sqrt(sum((validate_data$ejection_fraction-validate_data$pred_ejection_fraction)^2))



```

## Step 5b: Run basic Linear Regression using DICOM metadata to Predict Values from Test Dataset

```{r, echo=TRUE, message=FALSE, warning=FALSE,  fig.height = 8, fig.width = 8}

### Systolic Regression
sys_reg <- lm(Systole ~ patient_id + rows + columns + spacing + slice_thickness + plane + slice_count + up_down_agg + age_years + sex + small_slice_count + normal_slice_count + angle, data=train_data)
summary(sys_reg)

# Create Predicted Value
test_data$PredictedSystole <-predict(sys_reg, newdata = test_data)


# Plot Predicted vs Actual Systoliv Volume
ggplot(test_data, aes(x=test_data$Systole, y=test_data$PredictedSystole)) + 
  geom_point()+
  theme(plot.title = element_text(size=30)) + 
  labs(title="Predicted vs Actual Systolic Volume", x = "Predicted Volume (mL)", y = "Actual Volume (mL)") + 
  geom_smooth(method=lm)

# Calculate Key Metrics
cor(test_data$Systole, test_data$PredictedSystole)

# RMSE 
sqrt(sum((test_data$Systole-test_data$PredictedSystole)^2))
sqrt(sum((sys_reg$fitted.values-train_data$Systole)^2))

### Diastolic Regression

dias_reg <- glm(Diastole ~ patient_id + rows + columns + spacing + slice_thickness + plane + slice_count + up_down_agg + age_years + sex + small_slice_count + normal_slice_count + angle, data=train_data)
summary(dias_reg)

test_data$PredictedDiastole <-predict(dias_reg, newdata = test_data)

ggplot(test_data, aes(x=test_data$Diastole, y=test_data$PredictedDiastole)) + 
  geom_point()+
  labs(title="Predicted vs Actual Diastolic Volume", x = "Predicted Volume (mL)", y = "Actual Volume (mL)") + 
  theme(plot.title = element_text(size=30)) + 
  geom_smooth(method=lm)

cor(test_data$Diastole, test_data$PredictedDiastole)
sqrt(sum((test_data$Diastole-test_data$PredictedDiastole)^2))


### Ejection Fraction
ef_reg <- glm(ejection_fraction ~ patient_id + rows + columns + spacing + slice_thickness + plane + slice_count + up_down_agg + age_years + sex + small_slice_count + normal_slice_count + angle, data=train_data)
summary(dias_reg)

test_data$pred_ejection_fraction <-predict(ef_reg, newdata = test_data)

ggplot(test_data, aes(x=test_data$ejection_fraction, y=test_data$pred_ejection_fraction)) + 
  geom_point()+
  labs(title="Predicted vs Actual Ejection Fraction", x = "Predicted Ejection Fraction (%)", y = "Actual Ejection Fraction (%)") + 
  theme(plot.title = element_text(size=30)) + 
  geom_smooth(method=lm)

cor(test_data$ejection_fraction, test_data$pred_ejection_fraction)
sqrt(sum((test_data$ejection_fraction-test_data$pred_ejection_fraction)^2))




```

## Step 5: Calculate Key Metrics 

```{r, echo=TRUE, message=FALSE, warning=FALSE,  fig.height = 8, fig.width = 8}


```


